---
title: "Prevalence of Having Contraband among Pulled-over Drives"
author: "Yukun Li"
date: "12/4/2019"
output: pdf_document
---

# Background and Introduction
To be done.

# Causal Roadmap

## Scientific Question:
What is the prevalence of having contraband if all drives are searched.

## Causal Model
$W_1$ : age 

$W_2$ : race

$W_3$ : gender

$W_4$ : vehicle type

$\Delta$ : if search is conducted

Y* : Underlying contraband status

Y : if contraband is found

- Endogenous variables: $X = (W_1, W_2, W_3, W_4, \Delta, Y)$

- Exogenous variables: $U \sim \mathbb{P}_U$ (to be determined). 

Structral equation $F$:

\begin{align*}
W_1 &= f_{W1}(U_{W1})\\
W_2 &= f_{W2}(U_{W2})\\
W_3 &= f_{W3}(U_{W3})\\
W_4 &= f_{W4}(W_1, W_2, W_3, U_{W4})\\
\Delta &= f_\Delta(W_1, W_2, W_3, W_4, U_\Delta)\\
Y^* &= f_{Y^*}(W_1, W_2, W_3, W_4, U_{Y^*})\\
Y &= \Delta \times Y^*
\end{align*}

## Causal Parameter
$\Psi^*(\mathbb{P}^*) = \mathbb{P}^*(Y^* = 1) = \mathbb{P}^*(Y_{\Delta=1})$

## Observed data and its link to causal model
Observed data are randomly generated from the structual causal model.

## Identifiability
Lest's assume all $U$s are independent.

## Statistical estimand
$\Psi^*(\mathbb{P}^*)  = \Psi(\mathbb{P}_0) = \mathbb{E}_W\{\mathbb{P}_0(Y=1|\Delta=1,W)\}$

## Estimate
Parametric G-computation (simple substitution estimator), IPTW, TMLE. Use super learner during the estimating procedure. Don't forget to talk about the positivity assuptions.

Present a detailed plan for statistical inference/variance estimation based on the non-parametric bootstrap and implement it.

# Data preprocessing
```{r, message=FALSE}
# packages
library(tidyverse)
library(lubridate)
```

```{r}
# load dataset
dat <- readRDS("data/MAStatePatrol.rds")
# take a look at the variables we have
colnames(dat)
# the dataset is balanced over years, 
# we will use the observations only in 2015 for 
# computational convenience and intepretability of results.
table(year(dat$date))

dat_prep <- function(dat, loc, years){
  
  datBos <-  dat %>% 
  filter(location == loc) %>%
  filter(year(date) == years) %>% 
  filter(subject_race != 'unknown' & subject_race != 'other') %>% # positivity assumption
  filter(vehicle_type != 'Motorcycle' & vehicle_type != 'Trailer') %>% # positicity assumption
  filter(!is.na(subject_age) & !is.na(subject_sex)) %>% 
  select(subject_age,
         subject_race,
         subject_sex,
         vehicle_type,
         contraband_found,
         search_conducted,
         # the following variables are not used.
         outcome,
         frisk_performed,
         search_basis,
         reason_for_stop,
         raw_Race)
# Here I select all the varibales that might be useful. 
# A further discussion is needed to decide how to use them.

# drop unused levels from the dataframe
datBos <- droplevels(datBos)
return(datBos)
}

datBos <- dat_prep(dat, 'BOSTON', 2014)

# show summary and check postivity assumptions
summary(datBos$search_conducted)
summary(datBos$contraband_found)
summary(datBos$subject_age)
summary(datBos$subject_sex)
table(datBos$subject_race)
table(datBos$vehicle_type)

###############!!!!!###############
# positivity assumptions are heavily violated if we include vehicle type in W.
table(datBos$subject_race, datBos$subject_sex, datBos$vehicle_type, datBos$search_conducted)
table(datBos$subject_race, datBos$subject_sex, datBos$search_conducted)
```


Following are some attemps of estiamtion, I'll consider more parametric models and super learner and put all of them in R functions.

# G-computation
```{r}
# G-computation
# (1) NPMLE
# (2) logistic model : E(Y|D, W) ~ all Ws
# (3) may inlcude some interaction.
datBos.searched <-  datBos %>% filter(search_conducted == TRUE)
fit.gcomp <- glm(contraband_found ~ 
                   subject_age + 
                   as.factor(subject_race) + 
                   subject_sex + 
                   as.factor(vehicle_type),
                 family = 'binomial', data = datBos.searched)
datBos.intervene <- datBos %>% mutate(search_conducted = TRUE)
EY.gcomp <- predict(fit.gcomp, newdata = datBos.intervene, type = 'response')
est.gcomp <- mean(EY.gcomp)
est.gcomp
```

# IPTW
```{r}
# P0(A = 1|W) = logit^(-1){B0 + B1W1 + B2W2 + B2W3 + B4W4}
fit.prob.D <- glm(search_conducted ~ 
                    subject_age + 
                    as.factor(subject_race) + 
                    subject_sex + 
                    as.factor(vehicle_type),
                  family = 'binomial', data = datBos)
prob.D1 <- predict(fit.prob.D, type = 'response')
summary(prob.D1)
# calculate weights
wt1 <- as.numeric(datBos$search_conducted == 1)/prob.D1
summary(wt1)
est.IPTW <- mean(wt1*datBos$contraband_found, na.rm = TRUE)
est.IPTW # too large

# Stabelized IPTW
wt.mean <- mean(wt1[!is.na(datBos$contraband_found)])
est.sIPTW <- mean(wt1*datBos$contraband_found, na.rm = TRUE)/wt.mean
```

# TMLE
